---
title: "Multimodel"
author: "José Gabriel Escarramán"
date: "2024-04-23"
output:
  html_document:
    df_print: kable
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)
options(scipen = 999)
library(data.table)
library(tidyverse)
library(ggplot2)
library(dplyr)
#library(caret)
library(gmodels)
library(readxl)

#Survival
library(survminer)
library(survivalROC)
library(ggpubr)
library(survival)
library(sjPlot)

#Roc curve
library(cutpointr)
library(ROCR)
library(ROCit)
library(pROC)
library(timeROC)

# requested 
library(tidyr)
library(broom)
library(lattice)
library(kableExtra)
library(Hmisc)
```

## Multivariate Model


```{r cars, echo=FALSE, results='hide'}

#Cinzia
#setwd("C:/Users/vincenza.solli2/Alma Mater Studiorum Università di Bologna/MM group - Documenti/CARTELLE_PERSONALI/Escarraman_Jose/tesi/data//")
#jose
setwd("C:/Users/jose-/Alma Mater Studiorum Università di Bologna/MM group - Escarraman_Jose/tesi/data")

library(dplyr)

#final_data$ISS3 %>% table()
final_data <- read_excel("./Darafull_fu.xlsx", sheet = "data", na = "nv")
names2 <- read_excel("./Darafull_fu.xlsx",sheet="var", na = "nv")

cat<-names2$var[names2$type=="cat"]
#just for dara
#cat<- names2$var[names2$type=="cat" & names2$linea=="2"]



num<-names2$var[names2$type=="num"]
cat<- names2$var[names2$type=="cat"]



for (i in c(cat)) {
  print(i)
  final_data[[i]]<-as.factor(final_data[[i]])
}

# convert multiple colums to numeric
for (i in c(num)) {
  print(i)
  final_data[[i]]<-as.numeric(final_data[[i]])
}

final_data2<-final_data

final_data2 <- final_data2 %>% 
  filter(region_x !="Consenza")


#significative variables in the univariate analysis

OSv<-c("bestresp_DaraVd_m_SD"," bestresp_DaraVd_M_VGPR"," bestresp_DaraVd_M_CR"," bestresp_DaraVd","
     mg_settimanale_Dex_DaraVd","pd_BM_g60"," pd_PCR_g05" )

PFSv<-c("profilassi_antivirale_DaraVd"," bestresp_DaraVd_m_SD"," bestresp_DaraVd_M_VGPR"," bestresp_DaraVd_M_CR","
        bestresp_DaraVd","mg_settimanale_Dex_DaraVd","	EMD_PD"," amp1q_PD"," pd_BM_g60" )

TTPv<-c("profilassi_antivirale_DaraVd"," bestresp_DaraVd_m_SD"," bestresp_DaraVd_M_VGPR"," bestresp_DaraVd_M_CR","
        bestresp_DaraVd","	EMD_PD"," amp1q_PD","del17_PD"," pd_BM_g60")




```



```{r, echo=FALSE, results='hide'}


options(scipen = 999)

final_data2$PFS_MONTHS_daravd_linea2 <- as.numeric(final_data2$PFS_MONTHS_daravd_linea2)
final_data2$EVENT_PFS_DaraVd_linea2 <- as.character(final_data2$EVENT_PFS_DaraVd_linea2) %>%  as.numeric()



final_data2$OS_MONTHS_PD <- as.numeric(final_data2$OS_MONTHS_PD)
final_data2$OS_EVENT <- as.character(final_data2$OS_EVENT) %>%  as.numeric()


final_data2$TTP_DARAVD_months <- as.numeric(final_data2$TTP_DARAVD_months)
final_data2$EVENT_TTP_DaraVd_linea2 <- as.character(final_data2$EVENT_TTP_DaraVd_linea2) %>%  as.numeric()


#create temp functions for COX univariate
tmpfun_pfs <- function(x) as.formula(paste("PFS",x,sep="~"))
tmpfun_ttp <- function(x) as.formula(paste("TTP",x,sep="~"))
tmpfun_os <- function(x) as.formula(paste("OS",x,sep="~"))




PFS <- Surv(final_data2$PFS_MONTHS_daravd_linea2, final_data2$EVENT_PFS_DaraVd_linea2)
OS <- Surv(final_data2$OS_MONTHS_PD, final_data2$OS_EVENT)
TTP <- Surv(final_data2$TTP_DARAVD_months, final_data2$EVENT_TTP_DaraVd_linea2)


import<-final_data2

```


Per la multivariata avremmo pensato di fare un modello con:
-Best resp ≥VGPR (che sarebbe bestresp_DaraVd_M_VGPR, giusto?)

-EMD_PD0

-PD Clearance ≥ 40

-PD BM ≤ 60%

Poi, se si può, proverei un modello con stesse variabili + ampl1q alla PD  (5 variabili) e un altro con in più la del17 alla PD (6 variabili).




## Models for PFS
```{r, echo=FALSE}

library(sjPlot)
attach(final_data2)


MODEL_PFS1<-coxph(PFS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD +
                        pd_BM_g60+
                        pd_CLEARENCE_g40
                  )

summary(MODEL_PFS1)
library(sjPlot)

#plot_model(MODEL_PFS1, ci_level=0.95,exp=T)


##plot_model(MODEL_PFS1, type = "emm", terms = c("bestresp_DaraVd_M_VGPR", 
 #                       "EMD_PD"), show.data = T)
  #                      + 
   #                     pd_BM_g60+
    #                    "pd_CLEARENCE_g50" ), show.data = T)

#proportional hazard
#cz <- cox.zph(MODEL_PFS1)
#print(cz)
#plot(cz)


#survdiff(PFS ~ bestresp_DaraVd_M_VGPR, data = final_data2)




MODEL_PFS2<-coxph(PFS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60 +
                        pd_CLEARENCE_g40+
                        amp1q_PD)

summary(MODEL_PFS2)


##plot_model(MODEL_PFS2, ci_level=0.95,exp=T)


#proportional hazard

#cz <- cox.zph(MODEL_PFS2)
#print(cz)
#plot(cz)


MODEL_PFS3<-coxph(PFS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40+
                        amp1q_PD+
                        del17_PD
                        #pd_PCR_g05
                    )
 
MODEL_PFS4<-coxph(PFS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                       # pd_CLEARENCE_g40+
                        amp1q_PD+
                        del17_PD
                        #pd_PCR_g05
                    )

summary(MODEL_PFS3)

#plot_model(MODEL_PFS4, ci_level=0.95,exp=T)


 
#proportional hazard

#cz <- cox.zph(MODEL_PFS1)
#print(cz)
#plot(cz)




#
##plot_model(MODEL_PFS1, type = "emm", terms = c("bestresp_DaraVd_M_VGPR"+ 
            #            "EMD_PD"+ 
                        #pd_BM_g60+
             #           "pd_CLEARENCE_g50" ), show.data = T)
##plot_model(BEST_MODEL_PFS2, type = "emm", terms = c( "ISS", "cfDNAtumor_fraction_cut_off"), show.data = T)


##plot_model(BEST_MODEL_PFS2, type = "emm", terms = c( "ISS"), show.data = T)

#jugar con esto

##plot_model(MODEL_PFS1, type = "emm", terms = c("bestresp_DaraVd_M_VGPR", 
 #                       "EMD_PD"), show.data = T) #para chequiar la asociacion de dos vairiables y como impacta en el el efecto


#step(MODEL_PFS1, direction = "backward", trace=1, na.rm=TRUE)


```

## Models for OS
```{r include=FALSE, results='hide'}


attach(final_data2)
library(sjPlot)

MODEL_OS1<-coxph(OS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40)

summary(MODEL_OS1)

#plot_model(MODEL_OS1, ci_level=0.95,exp=T)


#proportional hazard
#cz <- cox.zph(MODEL_PFS1)
#print(cz)
#plot(cz)


#survdiff(PFS ~ bestresp_DaraVd_M_VGPR, data = final_data2)




MODEL_OS2<-coxph(OS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40+
                        amp1q_PD)

summary(MODEL_OS2)


#plot_model(MODEL_OS2, ci_level=0.95,exp=T)


#proportional hazard

#cz <- cox.zph(MODEL_PFS2)
#print(cz)
#plot(cz)


MODEL_OS3<-coxph(OS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40+
                        amp1q_PD+
                        del17_PD
                        #pd_PCR_g05
                    )
 


summary(MODEL_OS3)

MODEL_OS4<-coxph(OS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        #pd_CLEARENCE_g40+
                        amp1q_PD+
                        del17_PD
                        #pd_PCR_g05
                    )

 
#plot_model(MODEL_OS4, ci_level=0.95,exp=T, show.p = TRUE)


  
#proportional hazard

#cz <- cox.zph(MODEL_PFS1)
#print(cz)
#plot(cz)




#library(sjPlot)
##plot_model(MODEL_PFS1, type = "emm", terms = c("bestresp_DaraVd_M_VGPR"+ 
            #            "EMD_PD"+ 
                        #pd_BM_g60+
             #           "pd_CLEARENCE_g50" ), show.data = T)
##plot_model(BEST_MODEL_PFS2, type = "emm", terms = c( "ISS", "cfDNAtumor_fraction_cut_off"), show.data = T)


##plot_model(BEST_MODEL_PFS2, type = "emm", terms = c( "ISS"), show.data = T)



#step(MODEL_PFS1, direction = "backward", trace=1, na.rm=TRUE)


```


```{r, echo=FALSE}

summary(MODEL_OS1)

summary(MODEL_OS2)

summary(MODEL_OS3)

```

## PFS and OS curves strat by bestresp_DaraVd_M_VGPR

```{r, echo=FALSE}
PFSm1 <- survfit(PFS ~  final_data2$bestresp_DaraVd_M_VGPR, data = final_data2)
summary(PFSm1)


ggsurv <- ggsurvplot(
           PFSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

summary(coxph(PFS~bestresp_DaraVd_M_VGPR, data = final_data2))

separator<-"------------------------------------------OS-------------------------------------------"
separator

OSm1 <- survfit(OS ~  final_data2$bestresp_DaraVd_M_VGPR, data = final_data2)
summary(OSm1)


ggsurv <- ggsurvplot(
           OSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

summary(coxph(OS~bestresp_DaraVd_M_VGPR, data = final_data2))

```


```{r, echo=FALSE}

MODEL_PFS4<-coxph(PFS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40+
                        amp1q_PD+
                        tt(del17_PD)
                        #pd_PCR_g05
                    )


MODEL_OS4<-coxph(OS ~ 
                        bestresp_DaraVd_M_VGPR+ 
                        EMD_PD+ 
                        pd_BM_g60+
                        pd_CLEARENCE_g40+
                        amp1q_PD+
                        tt(del17_PD)
                        #pd_PCR_g05
                    )


```

## del17_PD
```{r, echo=FALSE}

amp1q_PD
                    

PFSm3 <- survfit(PFS ~  final_data2$del17_PD, data = final_data2)
summary(PFSm3)


ggsurv <- ggsurvplot(
           PFSm3,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv











OSm3 <- survfit(OS ~  final_data2$del17_PD, data = final_data2)
summary(OSm3)


ggsurv <- ggsurvplot(
           OSm3,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv



```



## amp1q_PD

```{r, echo=FALSE}

PFSm4 <- survfit(PFS ~  final_data2$amp1q_PD, data = final_data2)
summary(PFSm4)


ggsurv <- ggsurvplot(
           PFSm4,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv











OSm4 <- survfit(OS ~  final_data2$amp1q_PD, data = final_data2)
summary(OSm4)


ggsurv <- ggsurvplot(
           OSm4,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,70),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv



```