---
title: "Univariate analysis in PFS, TTP and OS"
author: "SV"
date: "2024-04-12"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)

options(scipen = 999)
library(data.table)
library(tidyverse)
library(ggplot2)
library(dplyr)
#library(caret)
library(gmodels)
library(readxl)

#Survival
library(survminer)
library(survivalROC)
library(ggpubr)
library(survival)

#Roc curve
library(cutpointr)
library(ROCR)
library(ROCit)
library(pROC)
library(timeROC)

# requested 
library(tidyr)
library(broom)
library(lattice)
library(kableExtra)
```

```{r cars, echo=FALSE, results='hide'}
#cinzia directory
#setwd("C:/Users/emat/Desktop/streamming/data/")
#biostat directory

#Cinzia
#setwd("C:/Users/vincenza.solli2/Alma Mater Studiorum Università di Bologna/MM group - Documenti/CARTELLE_PERSONALI/Escarraman_Jose/tesi/data//")
#jose
setwd("C:/Users/jose-/Alma Mater Studiorum Università di Bologna/MM group - Escarraman_Jose/tesi/data//")

#tempo_esposozione_lena_linea1


#final_data$ISS3 %>% table()
final_data <- read_excel("./Darafull_fu.xlsx", sheet = "data", na = "nv")
names2 <- read_excel("./Darafull_fu.xlsx",sheet="var", na = "nv")

cat<-names2$var[names2$type=="cat"]
#just for dara
#cat<- names2$var[names2$type=="cat" & names2$linea=="2"]



num<-names2$var[names2$type=="num"]
cat<- names2$var[names2$type=="cat"]



for (i in c(cat)) {
  print(i)
  final_data[[i]]<-as.factor(final_data[[i]])
}

# convert multiple colums to numeric
for (i in c(num)) {
  print(i)
  final_data[[i]]<-as.numeric(final_data[[i]])
}

final_data2<-final_data

#final_data2  <- final_data2 %>% 
#  filter(region_x !="Consenza")

```

```{r, echo=FALSE, results='hide'}


options(scipen = 999)

final_data2$PFS_MONTHS_daravd_linea2 <- as.numeric(final_data2$PFS_MONTHS_daravd_linea2)
final_data2$EVENT_PFS_DaraVd_linea2 <- as.character(final_data2$EVENT_PFS_DaraVd_linea2) %>%  as.numeric()



final_data2$OS_MONTHS_PD <- as.numeric(final_data2$OS_MONTHS_PD)
final_data2$OS_EVENT <- as.character(final_data2$OS_EVENT) %>%  as.numeric()


final_data2$TTP_DARAVD_months <- as.numeric(final_data2$TTP_DARAVD_months)
final_data2$EVENT_TTP_DaraVd_linea2 <- as.character(final_data2$EVENT_TTP_DaraVd_linea2) %>%  as.numeric()


#create temp functions for COX univariate
tmpfun_pfs <- function(x) as.formula(paste("PFS",x,sep="~"))
tmpfun_ttp <- function(x) as.formula(paste("TTP",x,sep="~"))
tmpfun_os <- function(x) as.formula(paste("OS",x,sep="~"))




PFS <- Surv(final_data2$PFS_MONTHS_daravd_linea2, final_data2$EVENT_PFS_DaraVd_linea2)
OS <- Surv(final_data2$OS_MONTHS_PD, final_data2$OS_EVENT)
TTP <- Surv(final_data2$TTP_DARAVD_months, final_data2$EVENT_TTP_DaraVd_linea2)


import<-final_data2

```

```{r, echo=FALSE, results='hide'}

#________dataframe output PFS_____________
output1<-output2<-NULL
for (i in cat) {
  if (sum(!is.na(final_data2[[i]]))>1) {
    if (length(levels(final_data2[[i]]))>=2) {
      print(i)
      out2<-coxph(tmpfun_pfs(i), data = import) %>% tidy( exponentiate =T, conf.int = T) %>% cbind(surv="PFS", .)
      
      output1<-rbind(out2)
      output2<-rbind( output1, output2)
    }
  }
}
output2

output<-output2[,c(1,2,3,7,8,6)]
names(output)<-c("PFS", "variable", "HR","Lower CI", "Upper CI", "p-value")
cox_univariates_PFS<-as.data.frame(cbind(output$PFS, output$variable, format(round(output$HR,3),nsmall = 3),  
                                     format(round(output$`Lower CI`,3),nsmall = 3),  format(round(output$`Upper CI`,3),nsmall = 3), format(round(output$`p-value`,4),nsmall = 3)))
names(cox_univariates_PFS)<-c("Time", "variable", "HR","Lower CI", "Upper CI", "p-value")


```

```{r, echo=FALSE, results='hide'}
#________dataframe output TTP_____________
output1<-output2<-NULL
for (i in cat) {
  if (sum(!is.na(final_data2[[i]]))>1) {
    if (length(levels(final_data2[[i]]))>=2) {
      print(i)
      out2<-coxph(tmpfun_ttp(i), data = import) %>% tidy( exponentiate =T, conf.int = T) %>% cbind(surv="TTP", .)

      output1<-rbind(out2)
      output2<-rbind( output1, output2)
    }
  }
}
output2

output<-output2[,c(1,2,3,7,8,6)]
names(output)<-c("TTP", "variable", "HR","Lower CI", "Upper CI", "p-value")
cox_univariates_TTP<-as.data.frame(cbind(output$TTP, output$variable, format(round(output$HR,3),nsmall = 3),  
                                     format(round(output$`Lower CI`,3),nsmall = 3),  format(round(output$`Upper CI`,3),nsmall = 3), format(round(output$`p-value`,4),nsmall = 3)))
names(cox_univariates_TTP)<-c("Time", "variable", "HR","Lower CI", "Upper CI", "p-value")


```

```{r, echo=FALSE, results='hide'}
#________dataframe output OS_____________
output1<-output2<-NULL
for (i in cat) {
  if (sum(!is.na(final_data2[[i]]))>1) {
    if (length(levels(final_data2[[i]]))>=2) {
      print(i)
      out2<-coxph(tmpfun_os(i), data = import) %>% tidy( exponentiate =T, conf.int = T) %>% cbind(surv="OS", .)
      
      output1<-rbind(out2)
      output2<-rbind( output1, output2)
    }
  }
}
output2

output<-output2[,c(1,2,3,7,8,6)]
names(output)<-c("OS", "variable", "HR","Lower CI", "Upper CI", "p-value")
cox_univariates_OS<-as.data.frame(cbind(output$OS, output$variable, format(round(output$HR,3),nsmall = 3),  
                                     format(round(output$`Lower CI`,3),nsmall = 3),  format(round(output$`Upper CI`,3),nsmall = 3), format(round(output$`p-value`,4),nsmall = 3)))
names(cox_univariates_OS)<-c("Time", "variable", "HR","Lower CI", "Upper CI", "p-value")


```

####### VARIABILI SIGNIFICATIVE IN PFS PER CREARE IL MODELLO MULTIVARIATO

```{r, echo=FALSE}


cox_univariates <- rbind(cox_univariates_PFS, cox_univariates_TTP, cox_univariates_OS)


cat2<- names2$var[names2$type=="cat" & names2$linea=="2"]
patterns <- paste(cat2, collapse = "|")


cox_univariates_PFS<- cox_univariates_PFS %>% 
filter(`p-value`<=0.05,
       grepl(patterns, variable)) %>% 
  as.data.frame()


kable(cox_univariates_PFS, caption = "Significative variables at 2nd line (DARA) for PFS") %>%
  kable_styling(font_size = 14) %>%
  column_spec(1:2, bold = TRUE)




```

####### VARIABILI SIGNIFICATIVE IN TTP PER CREARE IL MODELLO MULTIVARIATO

```{r, echo=FALSE}

cox_univariates_TTP<- cox_univariates_TTP %>% 
filter(`p-value`<=0.05,
       grepl(patterns, variable)) %>% 
  as.data.frame()

kable(cox_univariates_TTP, caption = "Significative variables at 2nd line (DARA) for TTP") %>%
  kable_styling(font_size = 14) %>%
  column_spec(1:2, bold = TRUE)


```

####### VARIABILI SIGNIFICATIVE IN OS PER CREARE IL MODELLO MULTIVARIATO

```{r, echo=FALSE}

cox_univariates_OS<- cox_univariates_OS %>% 
filter(`p-value`<=0.05,
       grepl(patterns, variable)) %>% 
  as.data.frame()

kable(cox_univariates_OS, caption = "Significative variables at 2nd line (DARA) for OS") %>%
  kable_styling(font_size = 14) %>%
  column_spec(1:2, bold = TRUE)

```

MEDIAN DARA PFS without strat

```{r, echo=FALSE}

PFSm <- survfit(PFS ~  1, data = final_data2)
summary(PFSm)

ggsurv <- ggsurvplot(
           PFSm,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv



```

MEDIAN DARA OS without strat

```{r, echo=FALSE}

OSm <- survfit(OS ~  1, data = final_data2)
summary(OSm)

ggsurv <- ggsurvplot(
           OSm,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv





```

MEDIAN DARA TTP without strat

```{r, echo=FALSE}

TTPm <- survfit(TTP ~  1, data = final_data2)
summary(TTPm)

ggsurv <- ggsurvplot(
           TTPm,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "TTP - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

```

MEDIAN DARA PFS- HR_esordio and HR_DARA

```{r, echo=FALSE}

PFSm1 <- survfit(PFS ~  final_data2$HR_ESORDIO, data = final_data2)
summary(PFSm1)


ggsurv <- ggsurvplot(
           PFSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

PFSm2 <- survfit(PFS ~  final_data2$HR_DARA, data = final_data2)

summary(PFSm2)

ggsurv <- ggsurvplot(
           PFSm2,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv



```

MEDIAN DARA OS- HR_esordio and HR_DARA

```{r, echo=FALSE}

OSm1 <- survfit(OS ~  final_data2$HR_ESORDIO, data = final_data2)
summary(OSm1)


ggsurv <- ggsurvplot(
           OSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

OSm2 <- survfit(OS ~  final_data2$HR_DARA, data = final_data2)

summary(OSm2)

ggsurv <- ggsurvplot(
           OSm2,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv



```

MEDIAN DARA PFS- ISS and R-ISS

```{r, echo=FALSE}

PFSm1 <- survfit(PFS ~  final_data2$ISS_stadio_esordio, data = final_data2)
summary(PFSm1)


ggsurv <- ggsurvplot(
           PFSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

PFSm2 <- survfit(PFS ~  final_data2$RISS_stadio_esordio, data = final_data2)

summary(PFSm2)

ggsurv <- ggsurvplot(
           PFSm2,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "PFS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
           risk.table.y.text.col = T,# colour risk table text annotations.
          #risk.table.height = 0.25, # the height of the risk table
           risk.table.y.text = F,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

```

MEDIAN DARA OS- ISS and R-ISS

```{r, echo=FALSE}

OSm1 <- survfit(OS ~  final_data2$ISS_stadio_esordio, data = final_data2)
summary(OSm1)


ggsurv <- ggsurvplot(
           OSm1,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv

OSm2 <- survfit(OS ~  final_data2$RISS_stadio_esordio, data = final_data2)

summary(OSm2)

ggsurv <- ggsurvplot(
           OSm2,                     # survfit object with calculated statistics.
           data = final_data2,             # data used to fit survival curves.
           risk.table = TRUE,
           surv.median.line = "hv",# show risk table.
           pval = F,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = c("Okabe-Ito"),
           xlim = c(0,60),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "OS - Time in months",   # customize X axis label.
           break.time.by = 6,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = T,# show bars instead of names in text annotations
                                    # in legend of risk table.
          #ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          #ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          #surv.median.line = "hv"  # add the median survival pointer.
          #legend.labs = c("Male", "Female")    # change legend labels.
        )

ggsurv





```

ALL univariate (without filter)

```{r, echo=FALSE}

cox_univariates <- cox_univariates %>%
  mutate(Significance = case_when(
    `p-value` < 0.01 ~ "III",
    `p-value` >= 0.01 & `p-value` < 0.05 ~ "II",
    `p-value` >= 0.05 & `p-value` < 0.1 ~ "I",
     `p-value` == "NA" ~ "",
    TRUE ~ ""
  )) %>%
  arrange(Time, `p-value`) %>%
  filter(variable!= "terapia_linea1Elo-RD" & variable!= "terapia_linea1RCD")

kable(cox_univariates, caption = "Significative variables univariate cox model for OS, PFS and TTP ") %>%
  kable_styling(font_size = 14) %>%
  column_spec(1:2, bold = TRUE)


```
